Blame and Coercion: Together Again for the Second Time

### 2019-10-18 paper.tex
clean up todo list

\todo[inline]{Big background}
Around 2006, many work on integrating dynamic typing and static typing emerge, 
notably gradual types \citep{siek2006gradual} and hybrid types 
\citep{flanagan2006hybrid}. The semantics of gradually typed languages is 
usuallly defined in terms of intermediate languages, where places for runtime 
type checking are explicitly stated in the forms of cast expressions. Examples 
of these intermediate languages include the blame calculus 
\citep{wadler2009well} 
and the cast calculus \citep{siek2009exploring}.
Representing casts naively, however, suffices sever space leak. 
For example, translation to intermediate languages can wrap some expressions 
with casts, including expressions at tail positions. Thus, mutually 
tail-recursive functions in source language can consume unbounded space at 
runtime. There are more examples of space leaking in \citet{herman2010space}.

\todo[inline]{The problem: space-efficient cast representation}
There are two best studied ways to represent cast space-efficiently: coercions 
\citep{herman2010space}\citep{siek2009exploring}\citep{siek2012interpretations} 
\citep{siek2015blame} and threesomes 
\citep{siek2010threesomes}\citep{garcia2013calculating}. They have different 
strength and 
weakness. \citet{garcia2013calculating} and \citet{siek2015blame} claim that 
coercion is easy to understand, but hard to implement, while threesome is easy 
to implement, but hard to understand. \citet{garcia2013calculating} in addition 
claims that coercion is more difficult to reason about formally because 
its compose function is not structurally recurisive. For instance, the 
developers of Grift \citep{kuhlenschmidt2018efficient} report that it is tricky 
to convince their proof assistants that the compose function terminates.

\todo[inline]{Our solution: hypercoercion}
We present a new cast representation, hypercoercion, which has a structurally 
recursive composition. When we prove theorems about hypercoercion in Agda, the 
definition ``just works''. 
Besides, We believe hypercoercion is about as easy to understand as coercion 
because their syntax definitions are similar. For instance, the 
coercion approach represents $ 
\cOOcast{\TOOdyn}{l}{\cOOcast{\POOunit}{l'}{\TOOdyn}} $ as $ \POOunit 
?^{l} ; \POOunit\! $, while the hypercoercion approach gives $ 
\hyperCoercionC{?^l}{\POOunit}{!} $. In Section~\ref{sec:background}, we review 
coercion together with cast calculus 
\citep{siek2009exploring}\citep{kuhlenschmidt2018efficient}. And we 
present hypercoerion in Section~\ref{sec:hypercoercion-definition}.

@@

\todo[inline]{The problem: proving correctness of cast representations}
Many of the existing research on coercions or threesomes include proof(s) of 
correctness, that is, the cast calculus reduces to an observation ($ e 
\Downarrow o $) if and only if their calcului reduce to the same observation ($ 
e \Downarrow_{*} o $), where an observation is either a value constructor or 
blame. 
Despite the striking similarity of these proofs, they are all built from ground 
up.

\todo[inline]{Our solution: a framework}
We present a framework for proving correctness of cast reprensentation 
(Section~\ref{sec:framework}), and use it to prove the correctness of 
hypercoercion (Section~\ref{sec:hypercoercion-correctness}). We are optimistic 
that this framework is general enough to include coercion and threesome.

@@

\subsection{The \lazyD{} cast calculus}
\todo[inline]{I want to merge this and the following subsetion to the 
subsection above. -KC}

@@

\todo[inline]{What is the main idea behind the organization of
 hypercoercions?}

@@

\todoKC{check consistency before using middle composition.}
$ c_1 \fatsemi^\ell c_2 = c' $ composes two hypercoercions. The target of $ 
c_1 $ and the source of $ c_2 $ might be different, in which case $ \ell $ must 
be a label.
$ m_1 \fatsemi^\ell (m_2,t) = (m',t') $ composes a middle with a 
pair of a middle and a tail. Similarly, the target of $ m_1 $ and the source of 
$ m_2 $ might be different, in which case $ \ell $ must be a label.
$ seq(c_1,c_2) $ composes two coercions where the target type of $ c_1 $ is 
equal to the source type of $ T_2 $.
$ id(T) $ constructs an identity coercion of $ T $ with the help of $ id(P) $.
$ cast(T_1,l,T_2) $ constructs a coercion from a source type, a label, and a 
target type.

@@

\todo[inline]{What is the framework? Why can it simply proofs? Soundness?}

@@

\todo[inline]{KC needs to describe added rules. -KC}
We list a fraction of reduction rules due to space limitation.
When values are constructed, their hypercoercion parts are filled with outputs 
of $ id $. For instance, when a function value is constructed, its first part 
and last part are initialized to $ id(T_1) $ and $ id(T_2) $ respectively.
When evaluating a cast expression, the current continuation is extended with a 
hypercoercion constructed by $ cast $. $ ext $ composes the new hypercoercion 
with the hypercoercion on the top of the continuation.
When evaluating a compound expression, the machine firstly construct the new 
pre-continuation, then turn it to a continuation by adding an identity 
hypercoercion at the top. For instance, when evaluating a \texttt{cons} 
expression, the machine firstly construct $ \hckOOconsI{e_2}{\rho}{\kappa} $, 
the new pre-continuation, then call $ cont $, which adds an identity cast to 
form a continuation. 
When a function call happens, the machine firstly cast the operand. If the 
casting succeeds, the machine then evaluate the body in the extended 
environment and the extended continuation. If the casting fails, the machine 
then halts with the blame label.

>>>>

Around 2006, many work on integrating dynamic typing and static typing emerge, 
notably gradual types \citep{siek2006gradual} and hybrid types 
\citep{flanagan2006hybrid}. The semantics of gradually typed languages is 
usuallly defined in terms of intermediate languages, where places for runtime 
type checking are explicitly stated in the forms of cast expressions. Examples 
of these intermediate languages include the blame calculus 
\citep{wadler2009well} 
and the cast calculus \citep{siek2009exploring}.
Representing casts naively, however, suffices sever space leak. 
For example, translation to intermediate languages can wrap some expressions 
with casts, including expressions at tail positions. Thus, mutually 
tail-recursive functions in source language can consume unbounded space at 
runtime. There are more examples of space leaking in \citet{herman2010space}.

There are two best studied ways to represent cast space-efficiently: coercions 
\citep{herman2010space}\citep{siek2009exploring}\citep{siek2012interpretations} 
\citep{siek2015blame} and threesomes 
\citep{siek2010threesomes}\citep{garcia2013calculating}. They have different 
strength and 
weakness. \citet{garcia2013calculating} and \citet{siek2015blame} claim that 
coercion is easy to understand, but hard to implement, while threesome is easy 
to implement, but hard to understand. \citet{garcia2013calculating} in addition 
claims that coercion is more difficult to reason about formally because 
its compose function is not structurally recurisive. For instance, the 
developers of Grift \citep{kuhlenschmidt2018efficient} report that it is tricky 
to convince their proof assistants that the compose function terminates.

We present a new cast representation, hypercoercion, which has a structurally 
recursive composition. When we prove theorems about hypercoercion in Agda, the 
definition ``just works''. 
Besides, We believe hypercoercion is about as easy to understand as coercion 
because their syntax definitions are similar. For instance, the 
coercion approach represents $ 
\cOOcast{\TOOdyn}{l}{\cOOcast{\POOunit}{l'}{\TOOdyn}} $ as $ \POOunit 
?^{l} ; \POOunit\! $, while the hypercoercion approach gives $ 
\hyperCoercionC{?^l}{\POOunit}{!} $. In Section~\ref{sec:background}, we review 
coercion together with cast calculus 
\citep{siek2009exploring}\citep{kuhlenschmidt2018efficient}. And we 
present hypercoerion in Section~\ref{sec:hypercoercion-definition}.

@@

Many of the existing research on coercions or threesomes include proof(s) of 
correctness, that is, the cast calculus reduces to an observation ($ e 
\Downarrow o $) if and only if their calcului reduce to the same observation ($ 
e \Downarrow_{*} o $), where an observation is either a value constructor or 
blame. 
Despite the striking similarity of these proofs, they are all built from ground 
up.

We present a framework for proving correctness of cast reprensentation 
(Section~\ref{sec:framework}), and use it to prove the correctness of 
hypercoercion (Section~\ref{sec:hypercoercion-correctness}). We are optimistic 
that this framework is general enough to include coercion and threesome.

@@

\subsection{The \lazyD{} cast calculus}

@@

\todo[inline]{What is the main idea behind the organization of
 hypercoercions? -JS
 \\---\\
 Are you asking why hc has three parts? Or why these subsections 
in this section? -KC}

@@

$ c_1 \fatsemi^\ell c_2 = c' $ composes two hypercoercions. The target of $ 
c_1 $ and the source of $ c_2 $ might be different, in which case $ \ell $ must 
be a label.
$ m_1 \fatsemi^\ell (m_2,t) = (m',t') $ composes a middle with a 
pair of a middle and a tail. Similarly, the target of $ m_1 $ and the source of 
$ m_2 $ might be different, in which case $ \ell $ must be a label.
$ seq(c_1,c_2) $ composes two coercions where the target type of $ c_1 $ is 
equal to the source type of $ T_2 $.
$ id(T) $ constructs an identity coercion of $ T $ with the help of $ id(P) $.
$ cast(T_1,l,T_2) $ constructs a coercion from a source type, a label, and a 
target type.

@@

@@

\todo[inline]{KC needs to add missing rules in and describe them. -KC}
We list a fraction of reduction rules due to space limitation.
When values are constructed, their hypercoercion parts are filled with outputs 
of $ id $. For instance, when a function value is constructed, its first part 
and last part are initialized to $ id(T_1) $ and $ id(T_2) $ respectively.
When evaluating a cast expression, the current continuation is extended with a 
hypercoercion constructed by $ cast $. $ ext $ composes the new hypercoercion 
with the hypercoercion on the top of the continuation.
When evaluating a compound expression, the machine firstly construct the new 
pre-continuation, then turn it to a continuation by adding an identity 
hypercoercion at the top. For instance, when evaluating a \texttt{cons} 
expression, the machine firstly construct $ \hckOOconsI{e_2}{\rho}{\kappa} $, 
the new pre-continuation, then call $ cont $, which adds an identity cast to 
form a continuation. 
When a function call happens, the machine firstly cast the operand. If the 
casting succeeds, the machine then evaluate the body in the extended 
environment and the extended continuation. If the casting fails, the machine 
then halts with the blame label.
